---
title: |  
  |
  |  Supplemental Information to:
  |
  |  Generalists link peaks in the shifting
  |  adaptive landscape of Australia's dragon lizards
  |
  |
  |
  |
  |
output:
  bookdown::pdf_document2:
      toc: false
      number_sections: false
      keep_tex: no
  #    pandoc_args: ["--lua-filter", "figure_caption_patch.lua"]
  #pdf_document: #default
  #    toc: false
  #    number_sections: false
  #word_document:
  #    reference_docx: /Users/ianbrennan/My Drive/R.Analyses/rMarkdown_CSLs/Word_Style_Reference_1.docx
fontsize: 12pt
linkcolor: orange 
geometry: margin = 1in
indent: true
csl: /Users/ianbrennan/My Drive/R.Analyses/rMarkdown_CSLs/current-biology.csl
bibliography: /Users/ianbrennan/Documents/GitHub/Tiliquini/Manuscript/Tiliquini_References.bib
linestretch: 1 # use to adjust line spacing
header-includes: 
- \usepackage{float}
- \usepackage[left]{lineno}
- \usepackage{graphicx,kantlipsum,setspace}
- \usepackage{caption}
- \captionsetup[figure]{font={stretch=1}}
- \captionsetup[table]{font={stretch=1}}
#- \linenumbers
#- \modulolinenumbers[5]
- \pagenumbering{gobble}
#- \pagenumbering{arabic}
- \newcommand{\beginsupplement}{  \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} # renames the supplemental figures as S...
- \usepackage{lscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \captionsetup[figure]{labelfont={bf},name={Figure},labelsep=period}

---

```{r echo=FALSE,warning=FALSE}
# this is so that figures are displayed in their proper place in the text, and not all at the end. 
library(knitr)
  opts_chunk$set(fig.path='figure/graphics-', cache.path='cache/graphics-', fig.align='center',
                 external=TRUE, echo=TRUE, warning=FALSE, fig.pos='H')
```

Document S1. Figures S1--S12 and Tables S1--S4

\newpage


```{=tex}
\beginsupplement
```


<!-- SUPPLEMENTARY FIGURES -->

# Supplementary Information

### Molecular Data
We used the *pipesnake* v1.2 workflow (Brennan et al. 2024) https://github.com/AusARG/pipesnake to handle, process, assemble, and align our molecular data, estimate gene trees and a species tree. All software employed for individual tasks are indicated in the Supplementary file software_versions.yml. 

In brief, the pipeline concatenates individual read files, then removes duplicate reads via BBMAP (Bushnell 2014), trims adapters and barcodes with TRIMMOMATIC (Bolger et al. 2014), identifies read pairs with PEAR (Zhang et al. 2014), removes off-target reads with BBMAP, assembles reads to contigs with SPAdes (Prjibelski et al. 2020), maps contigs to targets with BLAT (Kent 2002) and extracts the best hit for each target, carries out preliminary alignments with MAFFT (Katoh & Standley 2013), before refining the alignments with GBLOCKS (Talavera & Castresana 2007), estimates locus trees with IQ-TREE2 (Minh et al. 2020), and a species tree with weighted ASTRAL hybrid (Zhang & Mirarab 2022).

Assembled target sequences organized by sample (pseudo-reference genomes; PRGs) are available from XXX. Raw sequence data for newly sampled taxa are available from https://data.bioplatforms.com/dataset/?ext_search_by=&q=agamidae&organization=ausarg&sequence_data_type=Illumina+Capture&sort=score+desc%2C+metadata_modified+desc&ext_bbox= . 

### Optima Model Fitting
To explore the reliability of our results concerning the number and species assignment of optima we fit a number of alternative schemes. In addition to the base scheme of PCs 1-6 (>90% variance) and K=0-15, we also fit PCs 1-9 (>95% variance) and all 19 raw trait axes. To assess a more conservative view we also reduced the number of possible shifts, therefore limiting the number of optima to 5 (K=0-4). The results are highly consistent regardless of input data type (PCs vs. raw traits) or trait number (6, 9, 19), in identifying a core set of 4 shifts that are recovered in every scenario. Those correspond to (1) a baseline shift at the base of the Australian radiation of Amphibolurinae, and three shifts that correspond to primarily terrestrial taxa in (2) the edge leading to *Moloch*, (3) *Ctenophorus* + *Cryptagama*, and (4) *Tympanocryptis.* Additional shifts identified by favored (PCs 1-6) and less conservative (K>4) analyses identify optima that largely correspond to genera, including (5) *Lophosaurus*, (6) *Pogona* + *Rankinia*, (7) *Diporiphora*, and (8) the large-bodied clade of *Ctenophorus.* The remaining shifts identify outlying individual species that show unique morphologies, (9) *Diporiphora superba*, (10) *Chlamydosaurus kingii*, and (11) *Ctenophorus aurita*. The final regime belongs to the taxa at the base of the tree comprising *Intellagama* and *Hypsilurus*. 

### Sampling

The following Amphibolurinae taxa were unsampled in our molecular efforts:  
*Diporiphora convergens*\* STORR, 1974  
*Hypsilurus binotatus* (MEYER, 1874)  
*Hypsilurus bruijnii* (PETERS & DORIA, 1878)  
*Hypsilurus capreolatus* KRAUS & MYERS, 2012  
*Hypsilurus geelvinkianus* (PETERS & DORIA, 1878)  
*Hypsilurus godeffroyi* (PETERS, 1867)  
*Hypsilurus hikidanus* MANTHEY & DENZER, 2006  
*Hypsilurus longi* (MACLEAY, 1877)  
*Hypsilurus macrolepis* PETERS, 1872  
*Hypsilurus modestus* (MEYER, 1874)  
*Hypsilurus nigrigularis* (MEYER, 1874)  
*Hypsilurus ornatus* MANTHEY & DENZER, 2006  
*Hypsilurus papuensis* (MACLEAY, 1877)  
*Hypsilurus schultzewestrumi* (URBAN, 1999)  
*Hypsilurus tenuicephalus* MANTHEY & DENZER, 2006  
*Tympanocryptis uniformis*\* MITCHELL, 1948  

\*These species were described from a single specimen and have not been confirmed via genetic data (Melville & Wilson, 2019)

\  

To take advantage of all available morphological and phylogenetic information we incorporated three additional New Guinean *Hypsilurus* species into our dated tree used for trait evolution analyses. *Hypsilurus godeffroyi*, *H. papuensis*, and *H. nigrigularis* were added to our MCMCTree output using bind.tip in phytools (Revell 2024), based on the topology and ages estimated from a recent phylogenomic investigation of Squamata (Title & Singhal et al. 2024). 


### Node Priors and Agamids in the Fossil Record
Acrodontan lizards are known from the fossil record of Europe, Australia, and across Asia, with arguably the oldest described lineage originating in Myanmar (Wagner et al. 2021). Unfortunately, the oldest fossil taxa (*Protodraco*, *Gueragama*) are of uncertain phylogenetic placement and remaining fossils (*Uromastyx europaeus*, *Barbaturex*, *Tinosaurus*) are considerably younger than recent estimates of subfamily splits in agamids (Title et al. 2024; Burbrink et al. 2020). Agamid fossils are known from Australia (Covacevich 1990; Ramm 2025), however most are comparatively young and unlikely to be of value as calibrations. The exceptions are mid-Miocene agamid fossils from Riversleigh World Heritage Area assigned to *Sulcatidens* and ‘*Physignathus*’. The latter were ascribed to *Physignathus* when this genus applied to both *P. cocincinus* and *I. lesueurii*, and leaves the current placement of these fossils uncertain. They are clearly Australian in origin, and while similar to *Intellagama*, they also share affinities with *Physignathus*, *Hypsilurus*, and *Chelosania* (M. Hutchinson, *pers. comm.*). These fossils were found across a number of Riversleigh sites, all of which have been dated to 16-18 myo by Woodhead et al. (2016): Camel Sputum (17.8), Wayne’s Wok (17.8), Inabayance (17.8), Upper (17.8), and RSO (16.6). To use this fossil we applied it as a minimum on the crown of Australian Amphibolurinae, with a soft upper bound of 50 mya informed by estimates of the Amphibolurinae—Draconinae/Agaminae split as estimated by Burbrink et al. (2020) and Title et al. (2024). We applied a secondary calibration on the divergence between Amphibolurinae and Agaminae (Phrynocephalus) as a uniform prior with soft bounds (50-80 mya) following Burbrink et al. 2020 and Title et al. 2024, and another uniform prior with soft bounds (238-255 mya) on the crown of Lepidosauria representing the fossil *Sophineta*, arguably the oldest stem squamate. To assess the appropriateness of this calibration strategy we downsampled our species tree to a single representative of each genus and ran preliminary analyses in MCMCTree. 


### Calculating Morphological Distances
To estimate multivariate morphological distances we calculated euclidean distance between pairs of taxa. For Figure 4 this required identifying transitions among optima and measuring distances between ancestral taxa or between an ancestor and extant taxon (see function node.to.node in script XX_funspace.R). 


\newpage


# Supplementary Figures


```{r Concatenated, echo=FALSE, eval=T, out.height = '90%', fig.align="center", fig.cap="(ref:SF1)"}
knitr::include_graphics("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Fig.S1_MCMCTree_PriorPosterior_NodeLabels.pdf")
```
(ref:SF1) \small Plots of applied priors (blue), effective priors (pink), and posteriors (green) for each node in the tree indicate reasonable behavior. Node numbers correspond to the phylogeny plotted below, plotted with ape, where n129 is the root node. The effective prior (pink) shows how the interaction of multiple priors can shape the expected ages for a given node. This can be seen when comparing input (blue) and effective
(pink) priors on any calibrated node (e.g. n129, n130, n134).


\newpage

\blandscape

```{r FullMCMC, echo=FALSE, fig.pos='H', out.height = '85%', eval=T, fig.cap="(ref:SF2)"}
knitr::include_graphics("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Fig.S2_Animal_Measurements.pdf")
```
(ref:SF2) \small Schematic of morphological measurements taken and outlined in Table S3.

\elandscape

\newpage

```{r LocusInfo, echo=FALSE, fig.pos='H', eval=T, fig.cap="(ref:SF3)"}
knitr::include_graphics("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Fig.S3_Amphibolurinae_CorrPlot.pdf")
```

(ref:SF3) \small Plot of size-corrected trait correlations with traits grouped by the morphological module they correspond to (either head, body, limb, or tail). Blues indicate positive correlations, red indicate negative correlations, with saturation indicating increasing intensity. 

\newpage

```{r Animal, echo=FALSE, out.height = '78%', fig.pos='H', eval=T, fig.cap="(ref:SF4)"}
knitr::include_graphics("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Fig.S4_Amphibolurinae_Sampling.pdf")
```

(ref:SF4) \small  Locus type and number summarized by sample for newly sequenced samples. Dotted line indicates 50% of targeted sequences recovered. Samples in grey text at left were not included in final analyses. 

\newpage


```{r BEASTCI, echo=FALSE, eval=T, fig.align="center", out.height='100%', fig.cap="(ref:SF5)"}
knitr::include_graphics("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Fig.S5_Agamidae_hASTRAL_ASTRAL4.pdf")
```
(ref:SF5) \small Species tree from weighted ASTRAL-hybrid including all samples sequenced and agamid outgroups. Local posterior probabilities are 1 for all nodes except those indicated by an orange circle.

\newpage

```{r PriorPost, echo=FALSE, eval=T, fig.align="center", out.height='90%', fig.cap="(ref:SF6)"}
knitr::include_graphics("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Fig.S6_Amphibolurinae_Genera_gCF.pdf")
```
(ref:SF6) \small Phylogeny of the Amphibolurinae showing intergeneric relationships and support across individual loci. Gene concordance factors (gCF) show the proportion of gene trees which support a given bifurcation. Pie charts on branches indicate the percent of loci which support the presented bifurcation (gCF~1~, orange), one of each of the two other most common resolutions (gCF~2~, blue; gCF~3~, pink) or all other possibilities (gCF~p~, grey). Areas of low concordance (e.g. placement of *Lophosaurus*) indicate topological uncertainty, likely as a result of high levels of incomplete lineage sorting. 

\newpage

<!--\blandscape -->

```{r TraitVRTrees, echo=FALSE, eval=T, out.height = '90%', fig.align="center", fig.cap="(ref:SF7)"}
knitr::include_graphics("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Fig.S7_Trait_Variance.pdf")
```
(ref:SF7) \small Individual traits show different levels of variance among amphibolurine species, with the greatest variances occurring in size (geometric mean) and tail length. Some traits, such as hand length and upper arm length show vanishingly small variation once corrected for absolute size. (Right) Traits with higher variances are more likely to fit the fabric model, which better partitions large changes in trait values. 

\  

\  


```{r ModelFits, echo=FALSE, eval=T, fig.align="center", out.height='80%', fig.cap="(ref:SF8)"}
knitr::include_graphics("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Fig.S8_VTT_LTT.pdf")
```
(ref:SF8) \small The Amphibolurinae hypervolume expands early and late in their evolution, with a period of niche packing in the Miocene. Diversification and morphological expansion are coupled in the early radiation of dragon lizards, but are largely decoupled from the Miocene onwards. 

\newpage

```{r ElabInn, echo=FALSE, eval=T, fig.align="center", out.height='80%', fig.cap="(ref:SF9)"}
knitr::include_graphics("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Fig.S9_RF_PhyloEM_Optima5.pdf")
```
(ref:SF9) \small Adaptive peaks corresponding to morphological regimes identified under a multi-optima OU model in *PhyloEM* and applied to ancestral taxa using randomForests. (Bottom) Tree shows the preferred 5 regime model estimated from all 19 size corrected morphological traits, with regime shifts denoted by small white circles. (Top) Tree shows the preferred 5 regimes including likelihoods for ancestors as estimated by randomForest. The amphibolurine MRCA is estimated as a *Hypsilurus*/*Lophosaurus*-like arboreal dragon and the Australian MRCA is estimated as a semiarboreal generalist lizard (both indicated with black outline to pie).  

\newpage

```{r, echo=FALSE, eval=T, fig.align="center", out.height='80%', fig.cap="(ref:SF10)"}
knitr::include_graphics("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Fig.S10_RF_PhyloEM_Optima12.pdf")
```
(ref:SF10) \small Adaptive peaks corresponding to morphological regimes identified under a multi-optima OU model in *PhyloEM* and applied to ancestral taxa using randomForests. (Bottom) Tree shows the preferred 12 regime model estimated from the first 6 PC axes of the morphological data, with regime shifts denoted by small white circles. (Top) Tree shows the preferred 12 regimes including likelihoods for ancestors as estimated by randomForest. The amphibolurine MRCA is estimated as a *Hypsilurus*-like arboreal dragon and the Australian MRCA is estimated as a semiarboreal lizard (both indicated with black outline to pie).

\newpage

```{r, echo=FALSE, eval=T, fig.align="center", out.height='80%', fig.cap="(ref:SF11)"}
knitr::include_graphics("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Fig.S11_HeatMap_Euclidean_TreeOrder.pdf")
```
(ref:SF11) \small Heatmap of multivariate morphological euclidean distance among all pairs of amphibolurines. This presentation highlights the conservative evolution of *Tympanocryptis* along with the relative dissimilarity of novel species like *Chlamydosaurus* (more similar to *Lophosaurus*/*Hypsilurus*), *Ctenophorus aurita* (more similar to *Tympanocryptis*), and *Moloch.* Black boxes along the diagonal delineate genera or clades of morphologically similar taxa. 

\newpage

```{r, echo=FALSE, eval=T, fig.align="center", out.height='100%', fig.cap="(ref:SF12)"}
knitr::include_graphics("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Fig.S12_Distance2RegimeCentroid.pdf")
```
(ref:SF12) \small See caption below.


\small \textbf{Figure S12.} Multivariate Euclidean distance estimated between each extant species and estimated optima, as well as to the amphibolurine MRCA. Tree at bottom is colored according to the k = 5 model of morphological optima. Lollipop plots above show morphological distance between each extant species and the amphibolurine MRCA (black), primarily arboreal regime 1 (red), regime 2 (blue; *Tympanocryptis*), the primarily generalist regime 3 (yellow), regime 4 (green; *Moloch*), and regime 5 (orange; *Ctenophorus*). Based on this metric and our measured traits, the amphibolurine MRCA was most morphologically similar to *Intellagama*, *Physignathus*, and *Chlamydosaurus*. Terrestrial forms like *Ctenophorus aurita*, the *Ctenophorus adelaidensis* group, and *Tympanocryptis* are likely highly derived and morphologically distinct from the MRCA.

\newpage

<!-- SUPPLEMENTARY TABLES -->

# Software

All software and versions used for *pipesnake* V.12 assembly and analysis of molecular data.
```{bash, eval=F, size="tiny"}
ASSEMBLY_POSTPROCESSING:
  python: 3.8.3
ASTER:
  astral-hybrid: 1.16.3.4
BBMAP_DEDUPE: BBMAP_FILTER: BBMAP_REFORMAT:
  BBMAP - clumpify.sh: '39.01'
  BBMAP - reformat.sh: '39.01'
  BBMAP - bbmap.sh: ''
BLAT:
  blat: '36'
CUSTOM_DUMPSOFTWAREVERSIONS:
  python: 3.10.6
  yaml: '6.0'
GBLOCKS:
  gblocks: 0.91b
IQTREE:
  iqtree: 2.2.6
MAFFT:
  mafft: '7.221'
MAKE_PRG:
  python: 3.8.3
MERGE_TREES:
  BusyBox: v1.22.1
PARSE_BLAT_RESULTS:
  python: 3.8.3
PEAR:
  pear: 0.9.6)
PHYLOGENY_MAKE_ALIGNMENTS:
  python: 3.8.3
PREPARE_ADAPTOR:
  python: 3.8.3
QUALITY_2_ASSEMBLY:
  python: 3.8.3
SED:
  sed: This is not GNU sed version 4.0
SPADES:
  spades: 3.15.5
TRIMMOMATIC:
  trimmomatic: '0.39'
TRIMMOMATIC_CLEAN_PE:
  trimmomatic: '0.39'
Workflow:
  Nextflow: 23.04.4
  ausarg/pipesnake: '1.2'
```


\newpage


\blandscape

# Supplementary Tables

```{r, include=F, message=F, warning=F}
library(dplyr)
library(gt); 
library(kableExtra); options(knitr.kable.NA = '---')
library(googlesheets4)
gs4_deauth()
sampinfo <- read.csv("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Table.S1_Agamidae_SampleInfo.csv")
sampinfo <- dplyr::select(sampinfo, Family, Subfamily, Genus, Species, RegNo, AHE, gene, uce, total_loci, Source,
                          -filename, -Targets, -Order, -HigherOrder)
```


### Table S1. Molecular Sampling

```{r, echo=F, eval=T}
# Here's a more basic version
kbl(sampinfo, longtable=T, booktabs=T, linesep="") %>% 
#  add_header_above(c("Microhylidae"=6)) %>%
  kable_styling(latex_options=c("repeat_header"),
                font_size=6)
```


\elandscape

### Table S2. Fossil Calibrations

| Fossil Information	|   Calibration	|   Split/Position	|   Source |
|---------------------|---------------|-------------------|----------|
| Uniform—Sophineta	|   'B(2.38,2.55)'	|   Lepidosauria (Sphenodon + Squamata)	|   see below |
| Uniform—Secondary	|   'B(0.50,0.80)'	|   Amphibolurinae + (Agaminae + Draconinae)	|   Burbrink et al. 2020; Title et al. 2024 |
| Uniform—Riversleigh 'Physignathus'	|   'B(0.17,0.50)'	|   Hypsilurus + Remaining Australian Amphibolurinae	|   Covacevich 1990 |

*Note*. The root divergence between lepidosaurs (Sphenodon + squamates) is based on the fossil taxa Sophineta cracoviensis (Evans & Bialynicka, 2009), Megachirella wachtleri (Renesto & Posenato, 2003), and the Vellberg Jaw (Jones et al., 2013). These taxa represent stem squamates and rhynchocephalians, and so provide a soft lower bound on the crown divergence of Lepidosauria, with a soft upper bound provided by Protosaurus speneri (von Meyer, 1832).

<!-- \blandscape -->

\newpage

### Table S3. Morphological Measurements

| No. | Measurement         | Abbreviation |                                                 Method |
|-----|---------------------|-------------:|-------------------------------------------------------:|
| 1   | Snout-vent length   |          SVL |                 From the tip of the snout to the vent. |
| 2   | Snout-axilla length |          SAL |              From the tip of the snout to the midpoint |
|     |                     |              |                of the crease between the fore-limb and |
|     |                     |              |                       the body on the ventral surface. |
| 3   | Inter-limb length   |          ILL |                  Midpoint of the crease on the ventral |
|     |                     |              |                surface where the fore-limb connects to |
|     |                     |              |                the body, to the midpoint of the crease |
|     |                     |              |             on the ventral surface where the hind-limb |
|     |                     |              |                                  connects to the body. |
| 4   | Body width          |           BW |               From one lateral side of the body to the |
|     |                     |              |      other, where possible at the midpoint of the ILL. |
| 5   | Pelvic width        |           PW |                 From the midpoint of the crease on the |
|     |                     |              |               ventral surface where the left hind limb |
|     |                     |              |               connects to the body, to the midpoint of |
|     |                     |              |                the crease on the ventral surface where |
|     |                     |              |              the right hind limb connects to the body. |
| 6   | Pelvic height       |           PH |            From the top of the dorsal surface were the |
|     |                     |              |          PW was measured, to the bottom of the ventral |
|     |                     |              |                     surface where the PW was measured. |
| 7   | Head width          |           HW |                       Widest part of the head from one |
|     |                     |              |                 dorsal-lateral edge to the other edge. |
| 8   | Head length         |           HL |         From the nose tip, to the anterior of the ear. |
| 9   | Snout length        |           SN |     From the nasal opening to the anterior of the eye. |
| 10  | Eye diameter        |           ED |                 From one side of the eye to the other. |
| 11  | Head depth          |           HD |                From the top of the tallest part of the |
|     |                     |              |              head on the dorsal surface, to the bottom |
|     |                     |              |                  of the ventral surface under the jaw. |
| 12  | Tail width          |           TW |                         Measured at the vent, from one |
|     |                     |              |                 dorsal-lateral edge to the other edge. |
| 13  | Tail length         |           TL |         Measured from the vent to the tip of the tail. |
| 14  | Fore-limb length    |          FLL |              Measured fully extended from the midpoint |
|     |                     |              |             of the crease on the ventral surface where |
|     |                     |              |                   the front limb connects to the body, |
|     |                     |              |             to the end of longest toe (claw included). |
| 15  | Front foot          |        FFOOT |                From the base of the foot to the end of |
|     |                     |              |                       the longest toe (claw included). |
| 16  | Lower front limb    |          LFL |          Measured from the base of the lower fore limb |
|     |                     |              |   to the juncture where the limb meets the front foot. |
| 17  | Upper front limb    |          UFL |        Measured from the crease on the ventral surface |
|     |                     |              |              where the fore limb connects to the body, |
|     |                     |              |                    to the end of the lower front limb. |
| 18  | Hind-limb length    |          HLL |                 From the midpoint of the crease on the |
|     |                     |              |           ventral surface where the hind limb connects |
|     |                     |              | to the body, to the end of longest toe (claw included) |
| 19  | Hind foot           |        HFOOT |                  From the base of the hind foot to the |
|     |                     |              |                 end of the longest toe (claw included) |
| 20  | Lower hind limb     |          LHL |                Measured from the top of the knee joint |
|     |                     |              |              to the heel juncture where the limb meets |
|     |                     |              |                                        the front foot. |
| 21  | Upper hind limb     |          UHL |                Measured from the crease on the ventral |
|     |                     |              |                surface where the hind limb connects to |
|     |                     |              |                      the body, to the end of the knee. |

**Table S3. Morphological measurements collected (21).**

\newpage

| No. | Measurement      | Shorthand     |                                   Method |
|-----|------------------|:--------------|-----------------------------------------:|
| 3   | Interlimb length | Interlimb     |                             see Table S3 |
| 4   | Body width       | Body_Width    |                             see Table S3 |
| 5   | Pelvic width     | Pelvic_Width  |                             see Table S3 |
| 6   | Pelvic height    | Pelvic_Height |                             see Table S3 |
| 7   | Head width       | Head_Width    |                             see Table S3 |
| 9   | Snout length     | Snout_Eye     |                             see Table S3 |
| 10  | Eye diameter     | Eye_Diameter  |                             see Table S3 |
| 11  | Head depth       | Head_Depth    |                             see Table S3 |
| 12  | Tail width       | Tail_Width    |                             see Table S3 |
| 13  | Tail length      | Tail_Length   |                             see Table S3 |
| 17  | Upper front limb | Upper_Arm     |                             see Table S3 |
| 16  | Lower front limb | Lower_Arm     |                             see Table S3 |
| 15  | Front foot       | Hand          |                             see Table S3 |
| 21  | Upper hind limb  | Upper_Leg     |                             see Table S3 |
| 20  | Lower hind limb  | Lower_Leg     |                             see Table S3 |
| 19  | Hind foot        | Foot          |                             see Table S3 |
| 22  | Neck length      | Neck          |               Snout_Axilla - Head_Length |
| 23  | Posterior skull  | Pos_Skull     | Head_Length - (Snout_Eye + Eye_Diameter) |
| 24  | Pelvic gap       | Pelvic_Gap    |  Snout_Vent - (Interlimb + Snout_Axilla) |

**Table S3. Final morphological traits used for phenotypic analyses (19).**

\newpage

\blandscape

### Table S4. BayesTraits Model Fitting Results

```{r, echo=F, eval=T}
# Here's a more basic version
models <- read.csv("/Users/ianbrennan/Documents/GitHub/Amphibolurinae/Supplement/Table.S4_BayesTraits_MarginalLikelihoods.csv")
kbl(models, longtable=T, booktabs=T, linesep="") %>% 
#  add_header_above(c("Microhylidae"=6)) %>%
  kable_styling(latex_options=c("repeat_header"),
                font_size=6)
```

\elandscape

# Supplemental References

::: {#refs}

Bolger, A. M., Lohse, M., & Usadel, B. (2014). Trimmomatic: a flexible trimmer for Illumina sequence data. Bioinformatics, 30(15), 2114-2120.  

\  

Katoh, K., & Standley, D. M. (2013). MAFFT multiple sequence alignment software version 7: improvements in performance and usability. Molecular Biology and Evolution, 30(4), 772-780.  

\  

Kent, W. J. (2002). BLAT—the BLAST-like alignment tool. Genome Research, 12(4), 656-664.  

\  

Prjibelski, A., Antipov, D., Meleshko, D., Lapidus, A., & Korobeynikov, A. (2020). Using SPAdes de novo assembler. Current protocols in bioinformatics, 70(1), e102.  

\  

Talavera, G., & Castresana, J. (2007). Improvement of phylogenies after removing divergent and ambiguously aligned blocks from protein sequence alignments. Systematic biology, 56(4), 564-577.  

\  

Zhang, J., Kobert, K., Flouri, T., & Stamatakis, A. (2014). PEAR: a fast and accurate Illumina Paired-End reAd mergeR. Bioinformatics, 30(5), 614-620.

:::

\newpage


